Marlowe: financial contracts on blockchain

### 2018-05-17 tex/paper.tex
added some notes to the intro about what to include there, and what to do in the body too
Add more detailed explanations to embedded example

@@

\section{The model}
\label{section:model}

@@

\begin{haskellcode}
iCC1 :: IdentCC
iCC1 = IdentCC 1
\end{haskellcode}
And an identifier for the payment:
\begin{haskellcode}
iP1 :: IdentPay
iP1 = IdentPay 1
\end{haskellcode}
We can also create a sub-contract that allows the money to be redeemed by participant \haskellinline{1}:
\begin{haskellcode}
redeem_original :: Contract
redeem_original = RedeemCC iCC1 Null
\end{haskellcode}
Each participant has a say on who deserves the money: either participant \haskellinline{1} deserves a refund, which we 
represent with the number $0$; or participant \haskellinline{2} deserves a payment, which we represent with the number 
$1$. Because there is only one choice to make per participant, we use the same \haskellinline{IdentChoice} as their 
participant id. We can define a function that returns an observation that is true if and only if the participant 
\haskellinline{per} has chosen the number \haskellinline{c} as follows:
\begin{haskellcode}
chose :: Int -> ConcreteChoice -> Observation
chose per c = PersonChoseThis (IdentChoice per) per c
\end{haskellcode}
Then we can easily write a function that returns an observation that is true if and only if at least one of the 
participants \haskellinline{per} and \haskellinline{per'} has chosen the number \haskellinline{c} as follows:
\begin{haskellcode}
one_chose :: Person -> Person -> ConcreteChoice -> Observation
one_chose per per' val = (OrObs (chose per val) (chose per' val)) 
\end{haskellcode}
After that, we can write a function that returns an observation that is true if and only if at least 
two of the three participants \haskellinline{p1}, \haskellinline{p2}, and \haskellinline{p3} have chosen the number 
\haskellinline{c} as follows:
\begin{haskellcode}
two_chose :: Person -> Person -> Person -> ConcreteChoice -> Observation
two_chose p1 p2 p3 c = OrObs (AndObs (chose p1 c) (one_chose p2 p3 c))
                             (AndObs (chose p2 c) (chose p3 c))
\end{haskellcode}
Finally, we can write the escrow contract, thus:
\begin{haskellcode}
escrow :: Contract
escrow = CommitCash iCC1 1 (ConstMoney 450) 10 100
                    (When (OrObs (two_chose 1 2 3 0)
                                 (two_chose 1 2 3 1))
                          90
                          (Choice (two_chose 1 2 3 1)
                                  (Pay iP1 1 2 (AvailableMoney iCC1) 100
                                       redeem_original)
                                  redeem_original)
                          redeem_original)
                    Null
\end{haskellcode}

@@

The next primitive, \haskellinline{When}, waits for one of three things to happen:
\begin{enumerate}
 \item Two out of three choose $0$, for `\haskellinline{pay}'. 
 \item Two out of three choose $1$, for `\haskellinline{refund}'.
 \item The block 90 is added to the blockchain.
\end{enumerate}

If the third option happens first, the money is refunded. Otherwise, there is a \haskellinline{Choice} that will 
immediately refund the money to participant \haskellinline{1} unless two out of three chose option $1$; in the later 
case, \haskellinline{Pay} will give participant \haskellinline{2} the opportunity to claim the funds available in 
\haskellinline{iCC1} before block 100. If the funds are not claimed before block 100 they will also be refunded to 
participant \haskellinline{1}.

Reflecting on the example we can see that using Haskell definitions has made the contract substantially more 
comprehensible; it will also support more efficient operation, if we were to replace the repeated  expression 
\haskellinline{two_chose 1 2 3 1} above by a \haskellinline{where} clause, 
\begin{haskellcode}
escrow = ...
                    (When (OrObs (two_chose 1 2 3 0)
                                 chose_refund)
                          90
                          (Choice chose_refund
                           ...
         ...
         where 
         chose_refund = two_chose 1 2 3 1                 
\end{haskellcode}
then the repeated computation of the expression will be avoided.

>>>>

TO ADD
\begin{itemize}
\item 
Model vs implementation: add a section on the letter e.g. about distributed computation.
\item
In particular we're agnostic about UTxO and accounts, but want to be implementable on both.
\item
focus on some of the design decisions, but focus on the important aspects only and cut the stuff that's  more speculative/internal
\item  
\ldots
\end{itemize}

@@

\section{The Marlowe model}
\label{section:model}

@@

\begin{haskellcode}
iCC1 :: IdentCC
iCC1 = IdentCC 1
\end{haskellcode}
And an identifier for the payment:
\begin{haskellcode}
iP1 :: IdentPay
iP1 = IdentPay 1
\end{haskellcode}
We can also create a sub-contract that allows the money from the commitment with identifier \haskellinline{IdentCC 1} to 
be redeemed:
\begin{haskellcode}
redeem_original :: Contract
redeem_original = RedeemCC iCC1 Null
\end{haskellcode}
Once redeemed, the contract continues as \haskellinline{Null} since we expect this to be the last thing that is done.
Each participant has a say on who deserves the money: either participant \haskellinline{1} deserves a refund, which we 
represent with the number $0$; or participant \haskellinline{2} deserves a payment, which we represent with the number 
$1$. Because there is only one choice to make per participant, we use the same \haskellinline{IdentChoice} as their 
participant id. We can define a function that returns an observation that is true if and only if
\haskellinline{per} has chosen the number \haskellinline{c} for the choice \haskellinline{IdentChoice per} as follows:
\begin{haskellcode}
chose :: Int -> ConcreteChoice -> Observation
chose per c = PersonChoseThis (IdentChoice per) per c
\end{haskellcode}
Then we can easily write a function that returns an observation that is true if and only if at least one of the 
participants \haskellinline{per} and \haskellinline{per'} has chosen the number \haskellinline{c} as follows:
\begin{haskellcode}
one_chose :: Person -> Person -> ConcreteChoice -> Observation
one_chose per per' val = (OrObs (chose per val) (chose per' val)) 
\end{haskellcode}
Building on that, we can now write a function that returns an observation that is true if and only if at least 
two out of the three participants \haskellinline{p1}, \haskellinline{p2}, and \haskellinline{p3} have agreed in 
choosing the number \haskellinline{c} as follows:
\begin{haskellcode}
two_chose :: Person -> Person -> Person -> ConcreteChoice -> Observation
two_chose p1 p2 p3 c = OrObs (AndObs (chose p1 c) (one_chose p2 p3 c))
                             (AndObs (chose p2 c) (chose p3 c))
\end{haskellcode}
Finally, we can write the escrow contract, thus:
\begin{haskellcode}
escrow :: Contract
escrow = CommitCash iCC1 1 (ConstMoney 450) 10 100
                    (When (OrObs (two_chose 1 2 3 0)
                                 (two_chose 1 2 3 1))
                          90
                          (Choice (two_chose 1 2 3 1)
                                  (Pay iP1 1 2 (AvailableMoney iCC1) 100
                                       redeem_original)
                                  redeem_original)
                          redeem_original)
                    Null
\end{haskellcode}

@@

The next primitive, \haskellinline{When}, waits for one of three things to happen:
\begin{enumerate}
 \item The observation became true because two out of three have chosen $0$. 
 \item The observation became true because two out of three have chosen $1$.
 \item The observation remained false until 90 was published in the blockchain.
\end{enumerate}

If the third option happens, the money is refunded. Otherwise, there is a \haskellinline{Choice} that will 
immediately refund the money to participant \haskellinline{1} unless two out of three chose option $1$; in the later 
case, \haskellinline{Pay} will give participant \haskellinline{2} the opportunity to claim the funds available in 
the commitment with identifier \haskellinline{iCC1} before block 100 (by using the identifier \haskellinline{iP1} for 
the claim). If the funds are not claimed before block 100 they will also be refunded to 
participant \haskellinline{1}.

Reflecting on the example, we can see that using Haskell definitions has made the contract substantially more 
comprehensible; it will also support more efficient operation, if we were to replace the repeated  expression 
\haskellinline{two_chose 1 2 3 1} above by a \haskellinline{where} clause, 
\begin{haskellcode}
escrow = ...
                    (When (OrObs (two_chose 1 2 3 0)
                                 chose_refund)
                          90
                          (Choice chose_refund
                           ...
         ...
         where 
         chose_refund = two_chose 1 2 3 1                 
\end{haskellcode}
then the repeated computation of the expression will be avoided.
